I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-11-23 paper_io_sim.tex
Paper revised

		The data explosion era seen in recent years results in growing needs in 
		data storing, data analysis. 
		Large computing systems including grids, clusters, clouds, peer-to-peer, 
		volunteer computers have been increasingly used to solve the problem. 
		Despite of their powerful computing capability and technological improvements, 
		there are still bottlenecks in these systems. 
		When it comes to I/O, a bottleneck can be seen in the disk I/O due to hardware 
		limitations. Data caching using page cache is one of the solutions applied 
		to minimize disk accesses and improve the I/O performance. 
		In computing system and application optimization studies, one of the 
		effective approaches is using simulation tools to conduct experiments instead 
		of running applications on real infrastructures. 
		Although simulators are built with individual simulation model for each computing 
		resource, the majority does not take page cache into consideration, 
		which can lead to low simulation accuracy in many cases.

		In this paper, we propose a simulation model page cache to simulate I/O 
		with key features of page cache. 
		The model is build with the consideration to allow it to be integrated to 
		existing simulation tools. 
		We also implement the model with SimGrid using WRENCH. 
		Then we create some simulators for different use cases. 
		Finally, we analyze the results, compare them to results of the existing 
		tool as well as real execution.

@@

		In recent years, as a result of the growth of data generated from a variety 
		of data sources, we have seen increasingly growing demands in data storage, 
		processing and analysis. 
		Storing and processing such a massive amount of data require large 
		computing infrastructures. 
		Data processing is a complex process whose the performance is affected 
		by many factors and hindered by bottlenecks in the systems.
		One of the bottlenecks that should be taken into consideration is disk access 
		time due to hardware limitations. 
		Beside hardware technical improvements, architectural solutions have been 
		proposed and implemented in computing systems in order to tackle the 
		disk I/O problem. 
		Caching, a successful technique applied into memory to maximize the 
		I/O performance by keeping data in memory for future accesses,
		has been proving the effectiveness when employed in computer systems.

		Studying in optimizing not only the utilization of page cache but also other 
		computing resources is a need that has never been overlooked.
		The research on computing infrastructures to maximize performance, 
		shorten processing time, and minimize costs gets involved with experiments. 
		However, real platforms are complex and not built for the experimental 
		purposes due to limited and shared resources, operational costs, unstable 
		environments and the result reproducibility difficulties. 
		An alternative approach is to used simulation tools developed for research 
		and experimental purposes.
		The obvious advantages of using simulation are the freedom of choosing 
		simulated infrastructures, reproducible results and low costs.
		These simulators are developed and integrated with simulation models 
		of computing resources including CPU, network and storage.
		However, some reviews over many simulators report that there has been 
		a trade-off between accuracy and scalability due to the level of details in 
		simulation models \cite{casanova2014simgrid} \cite{byrne2017review}.
		Moreover, different simulators are developed by and for a specific community,
		which hinders them from being used in cross-domain.
		SimGrid is a simulation framework developed with the objectives of  
		balancing between accuracy and scalability, as well as versatility.

		In spite of the importance of page cache and the impacts it has on I/O 
		operations, its simulation model is absent in many simulators. 
		Some simulation tools try to model the page cache with cached data, 
		but their models lack of some key features \cite{nunez2012simcan} 
		\cite{nunez2012icancloud}. 
		Moreover, these simulators have a problem in scalability with 
		long simulation time regarding to their very detailed level of simulation.
		Some other works are able to capture some important features of page cache, 
		but they focus on very specific domains \cite{xu2018saving}.

		In this paper, we propose a simulation model of I/O with page cache.
		The model is integrated in WRENCH, an open-source library built 
		on top of SimGrid with more abstract API. 
		Then we evaluate our model with four simulators in four 
		experiment scenarios. 
		This paper is organized as follows. 
		In Section \ref{relatedwork} we provide details of page cache 
		existing simulation tools.
		In Section \ref{relatedwork} we propose our I/O with page 
		cache simulation model and how we implement it and our experiments 
		to evaluate the model.  
		Section \ref{results} compares the results of our model with the results 
		of original WRENCH and real execution results. 
		Finally, Section \ref{discussion} concludes and presents future 
		work.

@@

			In Linux, page cache is a part of memory, which includes pages referring to 
			physical pages on disk \cite{linuxdev3rd2010}. 
			The goal of the page cache is to avoid disk I/O by keeping data in memory 
			for later accesses, minimize I/O time since disk accesses are considerably 
			slower than memory accesses.
			When the kernel starts a read operation, it checks if required data is in page cache.
			If yes, data is read from memory. This is called a \textit{cache hit}. 
			Otherwise, data is read from disk and placed in page cache.
			For writing, page cache is implemented with three different strategies. 
			The first is \textit{direct~I/O} or \textit{no-write}, which means the page cache 
			is not involved in write operations. 
			The second is \textit{write-through}, in which data goes \textit{through} 
			page cache to disk is. In this case, data is written to both page cache and disk. 
			In the third strategy, which is \textit{write-back}, data are written directly 
			to disk as \textit{dirty} pages and these pages will be periodically flushed 
			to disk later by a \textit{writeback} process and then marked as not dirty. 
			The write-back strategy is considered to outperform write-through as well as 
			direct I/O since it delays disk writes and perform in batch later 
			\cite{linuxdev3rd2010}.

			In write-back, one of the keys is the cache eviction mechanism, which 
			removes pages from the page cache when available memory is insufficient. 
			The kernel chooses clean pages from page cache to evict. 
			If the clean pages are inadequate, dirty pages are flushed to disk, 
			marked as clean and available for eviction. 
			Different cache eviction algorithms have been developed and proposed
			\cite{owda2014comparison}.
			In Linux kernel, a \textit{two lists strategy (LRU/2)} is implemented based on 
			the least-recently-used (LRU) scheme, with an \textit{active list} only containing 
			pages accesses more than once, and an \textit{inactive list} containing 
			pages available for eviction. 
			Pages are moved from inactive list to active list if they are 
			accessed when residing in the inactive list. 
			The lists work on LRU manner, pages are added to the tail and removed 
			from the top. 
			The lists are also kept balanced by moving pages from the active list
			to the inactive list when the active list grows too large.

@@

			There has been the research and development of many simulation tools 
			and frameworks for cloud computing with the goal to simulate cloud, grid, 
			HPC platforms, etc. 
			A number of the most famous existing simulators have been analyzed 
			and compared in some reviews (e.g., \cite{byrne2017review}, \cite{casanova2014simgrid}). 
			These simulators provide simulation models of CPU, network, storage 
			with different approaches and levels of details.
			Some simulators are at packet-level, which allows to achieve 
			high simulation accuracy with fine-grain details such as storage simulation 
			in \textit{DiskSim} or network simulation in \textit{CloudSim} 
			and \textit{iCanCloud} \cite{casanova2014simgrid}. 
			However, the issues with the packet-level simulation are scalability due to 
			long simulation time when the simulated platforms and applications grow 
			to hundreds and thousands nodes. 
			SimGrid is an open source simulation platform that has been developed 
			not only to balance the trade-off between accuracy and scalability, 
			but also to achieve versatility, which allows the use of the framework 
			to simulate applications in multiple domains \cite{casanova2014simgrid}.

			However, I/O with page cache model has not been provided in SimGrid. 
			It only simulates storage with capacity and bandwidths.
			Lebre \textit{et al.} adds a storage simulation extension to SimGrid 
			to simulate file system,  but the page cache is not taken into account 
			\cite{lebre2015}. 
			SIMCAN is the simulation framework that models I/O with page cache 
			with data on disk can be kept in page cache \cite{nunez2012simcan}.
			Memory is also modeled in iCanCloud simulator with a component in charge 
			of managing memory accesses and cached data \cite{nunez2012icancloud}, 
			but one of the issues is scalability due to its sophisticated model.
			Ultimately, non of these simulators provide any cache write strategy, 
			cache eviction mechanism with dirty/clean data and the page cache LRU lists. 
			Xu \textit{et al.} applies four cache replacement policies to simulate in-memory 
			caching, but for the study in energy consumption of multi-tier heterogenous 
			network \cite{xu2018saving}.

			In this study, we chose SimGrid to implement our I/O with page cache model
			due to the balancing between accuracy and scalibility, the versatility allowing it 
			to be adopted in different domains, and the design enabling this 
			framework to be extended with new models. 
			The main issue users have to face when using SimGrid is everything has to be 
			implemented from scratch due to its low-level API. 
			To ease this challenge, WRENCH, an open-source library, is built on top 
			of SimGrid to abstract the low-level logic and expose a higher abstraction 
			level API \cite{casanova2018wrench}. 
			We integrate our model in WRENCH and use the SimGrid API to interact with 
			simulated storage resource.
			This approach allows us to leverage existing features in WRENCH and SimGrid.
			The first is the process control, synchronization of resource consumption by 
			concurrent processes handled by the simulation core in SimGrid. 
			The second is the existing SimGrid storage model with capacity and 
			bandwidth properties, which are similar to the characteristics of memory.
			The third is dropping the burden of implementation with the high level 
			of abstraction in WRENCH.

@@

		In this section, we present our approach to model file read and write with
		I/O to page cache, cache eviction and data flushing mechanisms implemented 
		in the Linux kernel. We also detail the design of our simulators and their
		implementation. Finally, we describe experimental scenarios to evaluate
		our simulation model on real applications using a standalone Python
		prototype, SimGrid framework using original WRENCH and the WRENCH version 
		that is integrated with our simulation model (WRENCH-Ext).

@@

            We separate our simulation model in two components, the IO
			Controller and the Memory Manager, which together simulate 
			file read/write mechanisms in Linux (Figure~\ref{fig:interaction}). 
			The Memory Manager simulates I/Os to cache, flushing, periodical flushing, 
			and cache eviction, whereas the IO Controller simulates file reads and writes. 
			Simulated applications can simply interact with the IO Controller instead of 
			calling simulated storage directly as is the case with current simulators.

			\begin{figure}
   				\centering
   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
   				\caption{Overview of the simulated write-back page cache}\label{fig:interaction}
			\end{figure}

@@

			We leverage existing storage simulation models to simulate the memory,
			which is characterized with storage capacity, distinguished read/write 
			bandwidths and device latency. 
			All of these features are dealt within models in simulation 
			frameworks, which makes storage models applicable to memory \cite{lebre2015}. 
			In addition, leveraging existing disk models allows us 
			to take advantage of bandwidth sharing between concurrent disk I/O to 
			simulate parallel memory accesses.

			Memory Manager simulates two parallel threads. The first thread is the 
			main thread where activities of applications are simulated, and responsible 
			for the simulation of data flushing, cache eviction and I/Os to cache. 
			The second thread is a simulated background thread, in which Memory Manager 
			periodically searches and flushes expired dirty data in page cache LRU lists 
			to disks.

			In Linux kernel, the page cache LRU lists contain file pages. However, 
			due to the large number of file pages, managing lists of pages may 
			require significant computation overhead. 
			Thus, we introduce the concept of a data block as a unit to represent data 
			cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
			A data block has information about file name, block size, last access 
			time, a dirty flag that represents whether the data is clean (0) 
			or dirty (1), and the time it is written to cache as dirty data. 
			Blocks can have different sizes and a given file can have multiple 
			data blocks in page cache. In addition, a data block can be split into an 
			arbitrary number of smaller blocks.

@@

			Now, page cache LRU lists can be modeled with two lists of data blocks: 
			an active list and an inactive list, ordered by last access time 
			(earliest first, see Figure~\ref{fig:lrulist}).
			As in the kernel, our simulator keeps the size of the active list under
			twice the size of the inactive list by moving least recently 
            used data blocks from the active list to the inactive list.
			Modeling page cache LRU lists as lists of data blocks reduces the
			overhead of the simulator, as data blocks are obviously less
			numerous than file pages, while achieving the accuracy in I/O time
			simulation that can be seen in the later section.

			Given an arbitrary file, at a specific time, it can be partially cached, 
			completely cached, or uncached.
			Cached data is composed of data blocks existing in either or both
            of the page cache LRU lists. Cached blocks reside in the inactive list
            unless they have been recently reaccessed. Cached blocks 
            located on the inactive list that are subsequently accessed, 
            will be moved to the top of the active list. Cached blocks 
            recently written to cache are dirty blocks until they are flushed to disk.

            Our model supports chunk-by-chunk file read/write simulation
            with an user-defined chunk size, in which chunks are read/written 
            until file is entirely read/written. 
            The simulated LRU lists are used as follows. 
            When a file is read, its data can be partially cached, completely cached, 
            or uncached. 
            Uncached data is read from disk first, and cached data is only read 
            after all uncached data is read (left to right in Figure~\ref{fig:read_order}).
            When a chunk of uncached data is read, a new clean block is created 
            and put to the top of the inactive list. 
            For a chunk of cached data, because one or more existing data blocks 
            in two page cache LRU lists are accessed again, read dirty blocks are 
            moved to the active list, while clean data blocks are merged into 
            a single clean block and then also put in the active list. 
			The reason we do not merge dirty blocks is that different blocks 
			are written to cache at different time, and this time is used later for 
			periodical flushing process.
			In case a block is not entirely read, it is split into two smaller blocks, 
			one is reaccessed and the other remains the same.

			\begin{figure}
   				\centering
   				\includegraphics[width=\columnwidth]{figures/read_order.pdf}
   				\caption{File data read order}	\label{fig:read_order}
			\end{figure}

			In the case of writing data, we can assume that all data to be written is 
			uncached and it is written to cache as dirty data. 
			Thus, each time a chunk is written, we create a block of dirty data and 
			put it in the inactive list.

			\textcolor{red}{[Is there any kind of accesses other than read and write?]}.

@@

			Next, we simulate data flushing, which runs on the main simulated 
			thread and writes back dirty data to disk. 
			The data flushing simulation function traverses the sorted 
			inactive list first, and then the active list, looks for least recently used 
			dirty blocks, writes them to disk through the storage device model, 
			and sets dirty flags of these blocks to 0 until the total flushed amount 
			reaches the amount to be flushed or there is no dirty data left in cache. 
			If the last flushed block is not entirely flushed, it is split into 
			two blocks, one that is flushed and one that remains dirty.
			The flushing time is simulated with storage model.

			Similarly to data flushing, our cache eviction simulation runs on 
			the main thread, frees up the page cache by traversing and deleting 
			least recently used clean data blocks in the inactive list.
			The amount of evicted data is passed in and data blocks are deleted 
			from the inactive list until total evicted data reaches the amount 
			passed in, or there is no clean block left in the list.
			If the last evicted block is not entirely evicted, the block is split, 
			and only one block is deleted.
			The cache eviction simulation does not add up to the simulated time 
			since cache eviction time is negligible in real systems.

			\begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{in}{page cache inactive list}
        				\Desc{ac}{page cache active list}
						\Desc{t}{predefined flushing time interval}
						\Desc{sm}{storage simulation model}
   					\EndInput
   					\While{host is on}
						\State blocks = expired dirty blocks in active and inactive list
						\State flushing\_time = sm.write(blocks)
						\State Flag blocks as clean data
						\If{flushing\_time $<$ t}
							\State sleep(t - flushing\_time)
						\EndIf
					\EndWhile
				\end{algorithmic}
			\end{algorithm}

			The periodical flushing simulation is run on a simulated background thread 
			to flush expired dirty data blocks from cache to disk.
			Similar to expired dirty pages in real Linux systems, a dirty block in our model 
			is considered expired if the time it remains in cache as dirty is longer than 
			the predefined expire time. This time is counted from the time it is written 
			to cache to current simulated time. 
			The simulation algorithm of periodical flushing is an infinite loop in which 
			Memory Manager searches and flushes these expired dirty blocks to disk, 
			as described in Algorithm~\ref{alg:pdflush}. 
			The loop ends when the host the host is off (line 6). 
			In each repetition, Memory Manager finds expired dirty blocks in two 
			page cache LRU lists (line 7), simulates writes of data of these blocks 
			to disk (line 8), and mark them as clean (line 9).
			If the flushing time does not exceed our time interval, the thread is put 
			to sleep for the remaining time (lines 10-12). 
			Then based on the state of the host at current simulated time, 
			the algorithm continues or finishes the loop .
			Because periodical flushing is simulated as a background thread, it can be 
			concurrent with disk I/O on main thread. This parallelism is handled by the 
			storage model and reflected in simulated I/O time.

@@

			When simulated applications need to read/write a file, they can simply call 
			the IO Controller instead of the storage simulation models.

			As mentioned, in our model, a file read/write can be completed by read/write 
			chunk-by-chunk until entire file is read/written.
			To read or write a chunk of file, a simulated application sends a request to the 
			IOController instance on the node where the requested file is stored.
			Based on the memory status in Memory Manager, the IOController 
			orchestrates cache and disk reads/writes, flushing, cache eviction 
			to simulate I/O time.

>>>>

		The explosive development of Big Data era seen in recent years results in 
		growing needs in data storing, data analysis. 
		Although large computing systems with powerful computing capability 
		have been increasingly used to process big datasets, there is still a bottleneck 
		in disk I/O. 
		Memory page cache is a solution applied to minimize disk accesses and 
		improve the I/O performance. 
		Despite of the fact that simulation tools for HPC have been developed to 
		support studies in computing systems optimization, page cache simulation 
		has not been fully captured, or absent in these simulators. 
		The lack of ability to simulate I/O with page cache leads to inaccurate 
		simulation results in many situations, especially in data intensive applications.

		In this paper, we propose a I/O simulation model with key features of the page cache.
		The model consists of simulation algorithms of I/O, main components 
		and mechanisms of page cache.
		We also implement the model with SimGrid, a simulation framework for 
		distributed systems, using WRENCH, a SimGrid-based workflow simulator. 
		Then we create simulators for applications in single-threaded, multi-threaded,
		network file system and neuroimaging usecases.
		Finally, we analyze the results, compare to the results of base-line WRENCH 
		simulator as well as real execution.
		The results show that our simulator not only achieves higher accuracy than 
		original WRENCH, but also simulates well the status of memory and 
		page cache during I/O operations. 
		Through this study, our effort is to contribute an analytical model 
		of the page cache, which can also be integrated to other simulators 
		to improve I/O simulation accuracy, with its practical results.

@@

		In recent years, as a result of the growing demand for Big Data processing 
		on large computing infrastructures, there have been the research and development 
		of simulation applications and frameworks for distributed systems. 
		These simulation tools have shown their advantages in evaluation 
		of algorithms, applications and solutions. 
		Experimenting on real platforms exposes shortcomings due to limited and 
		shared resources, operational costs, unstable environments and the result 
		reproducibility difficulties.
		On the contrary, simulators provide to researchers the freedom of choosing 
		simulated infrastructures, reproducible results and low costs.
		These simulators are developed and integrated with simulation models 
		of computing resources including CPU, network and storage.
		The ultimate goal that all simulators try to achieve is to accurately simulate 
		applications. 
		However, there is a trade-off between accuracy and scalability, which has been 
		indicated in some reviews \cite{casanova2014simgrid} \cite{byrne2017review}. 
		This means that in most ciscumstances, users have to, more or less, sacrifice 
		accuracy for the scalability and performance when their platforms grow to 
		hundreds or thousands nodes.

		Beside computing resources such as CPU, network, storage, another 
		important factor that has significant impacts on I/O performance is 
		page cache. 
		Page cache is an architectural solution employed in systems to tackle 
		the problem of disk I/O bottleneck in data processing. 
		Although this is not a radical solution for disk performance, it contributes 
		to minimizing disk accesses and maximizing I/O performance. 
		Surprisingly, page cache has not been fully modeled, if not completely absent, 
		in simulation tools. 
		Some simulators try to simulate I/O with page cache, but fail to capture the 
		key features such as dirty data and cache eviction mechanism 
		\cite{nunez2012simcan} \cite{nunez2012icancloud}.
		Again, these simulators also have a problem in scalability with 
		long simulation time regarding to their very detailed level of simulation. 
		Other works are able to capture important features of page cache, 
		but they focus on very specific domains, such as energy consumption 
		in \cite{xu2018saving}.

		In this work, we propose a simulation model of I/O with page cache, 
		implement it with SimGrid, a simulation framework for distributed computing, 
		and WRENCH, an open-source library for developing workflow simulator 
		built on top of SimGrid. 
		One of the reasons for us to choose SimGrid is that it has a balance 
		between accuracy and scalability. Moreover, the versatility of SimGrid 
		enables it to be used in cross-domain \cite{casanova2014simgrid}. 
		In addition, SimGrid has been actively maintained with updated documents. 
		Then we evaluate our model with four simulators in four experiment scenarios. 
		This paper is organized as follows. 
		In Section \ref{relatedwork} we provide details of page cache 
		existing simulation tools.
		In Section \ref{relatedwork} we propose our I/O with page 
		cache simulation model and how we implement it and our experiments 
		to evaluate the model.  
		Section \ref{results} compares the results of our model with the results 
		of original WRENCH and real execution results. 
		Finally, Section \ref{discussion} concludes and presents future 
		work.

@@

			Linux implements a disk cache called \textit{page cache}. 
			It contains pages referring to physical pages on disk \cite{linuxdev3rd2010}.
			The goal of the page cache is to avoid disk I/O by keeping data in memory 
			for later accesses, minimize I/O time since disk accesses are considerably 
			slower than memory accesses.
			When the kernel starts a read operation, it checks if required data is in page cache.
			If yes, data is read from memory. This is called a \textit{cache hit}. 
			Otherwise, data is read from disk and placed in page cache.
			For writing, page cache is implemented with three different strategies. 
			The first is \textit{direct~I/O} or \textit{no-write}, which means the page cache 
			is not involved in write operations. 
			The second is \textit{write-through}, in which data goes \textit{through} 
			page cache to disk. In this case, data is written to both page cache and disk. 
			In the third strategy, called \textit{writeback}, data are written directly 
			to disk as \textit{dirty} pages and these pages will be periodically flushed 
			to disk later by a \textit{writeback} process and then marked as not dirty. 
			The writeback strategy is considered to outperform write-through as well as 
			direct I/O as it delays disk writes and perform a bulk write later.
			\cite{linuxdev3rd2010}.

			In writeback, one of the keys is the cache eviction mechanism, which is
			triggered when available memory is insufficient to removes pages 
			from page cache. 
			Because dirty pages are data written to page cache but yet to be 
			written to disk, dirty pages must be kept in page cache and only clean pages 
			are available for eviction.
			If the clean pages are inadequate, dirty pages are flushed to disk, 
			marked as clean and available for eviction. 
			There is another type of flushing mechanism, which is 
			\textit{periodical flushing}, called periodically in predefined intervals. 
			Periodical flushing only flushes expired dirty pages, which remain dirty in 
			page cache longer than an expiration time configured in the kernel. 
			Different cache eviction algorithms have also been developed and proposed
			\cite{owda2014comparison}.
			In Linux kernel, a \textit{two lists strategy (LRU/2)} is implemented based on 
			the least-recently-used (LRU) scheme, with an \textit{active list} and 
			an \textit{inactive list}. 
			If accessed pages are not in the page cache, they are added to the inactive list. 
			If pages are accessed when residing in the inactive list, they will be moved from 
			inactive list to active list. 
			The lists are also kept balanced by moving pages from the active list
			to the inactive list when the active list grows too large.
			Thus, the active list only contains pages which are accessed more than once 
			and not evictable, while the inactive list includes pages accessed once only, 
			or pages accessed more than once but moved from the active list. 
			The lists work in LRU manner, pages are added to the tail and removed 
			from the top.

@@

			Many simulation frameworks and applications have been developed 
			in recent years to support studies in grids, cloud, HPC, peer-to-peer 
			networks, volunteer computers. 					
			These simulators are built with models of resources such as CPU, network, 
			storage with different approaches and levels of details in different simulators. 
			The detail level of the model causes a conflict between accuracy and scalability, 
			and these simulators usually trade of one for the other 
			\cite{casanova2014simgrid}. 
			For example, some simulators are built with packet-level network models 
			such as \textit{CloudSim} and \textit{iCanCloud}, or storage model that 
			simulates the operation of storage hardware in \textit{DiskSim} 
			\cite{casanova2014simgrid}. 
			Fine-grained models allow simulators to achieve high accuracy, 
			but at the same time lead to scalability issues due to long simulation time 
			when simulated platforms and applications grow. 
			SimGrid is an open source simulation platform that has been developed 
			not only to balance the trade-off between accuracy and scalability, 
			but also to achieve versatility, which allows the use of the framework 
			to simulate applications in multiple domains \cite{casanova2014simgrid}.

			Despite the fact that many sophisticated models for resources have been 
			integrated in simulators, and the page cache shows apparent impacts on I/O
			performance, the simulation model of page cache is still missing in most 
			of simulators.
			For example, SimGrid only simulates I/O operations with storage bandwidths 
			and capacity. 
			SIMCAN is a simulation framework that models page cache by storing data 
			accessed on disk in a block cache \cite{nunez2012simcan}. 
			Page cache is also modeled in iCanCloud simulator with a component in charge 
			of managing memory accesses and cached data \cite{nunez2012icancloud}, 
			but one of the issues is scalability due to its sophisticated model.
			Ultimately, non of these simulators provide any cache write strategy, 
			cache eviction mechanism with dirty/clean data and the page cache LRU lists. 
			Cache replacement policies is applied in \cite{xu2018saving} to simulate 
			in-memory caching, but for the study in energy consumption of multi-tier 
			heterogenous network.

			In this study, we chose SimGrid to implement our model of I/O with page cache 
			due to the balancing between accuracy and scalibility, the versatility allowing it 
			to be adopted in different domains. 
			In addition, it is an actively maintained open-scource framework with 
			modular design, which enables the framework to be extended with new models. 
			The main issue users have to face when using SimGrid is everything has to be 
			implemented from scratch due to its low-level API. 
			To ease this challenge, WRENCH, another open-source library, is built on top 
			of SimGrid to abstract the low-level logic and expose a higher abstraction 
			level API \cite{casanova2018wrench}. 
			We integrate our model in WRENCH and use the SimGrid API to interact with 
			simulated storage resource.
			This approach allows us to leverage existing features in WRENCH and SimGrid.
			Firstly, we can take advantage of the process control, synchronization of 
			resource consumption by concurrent processes provided by the simulation core 
			in SimGrid. 
			Secondly, we can reuse the existing SimGrid storage model with capacity 
			and bandwidth properties, which are similar to the characteristics of memory.
			And finally, WRENCH provides a high level abstraction, which help us drop 
			the burden of implementation.

@@

		In this section, we present our approach to model file read and write with 
		page cache, cache eviction and data flushing mechanisms implemented 
		in the Linux kernel. 
		We also detail the design of our simulator and its implementation. 
		Because the simulator is implemented with SimGrid using WRENCH, we 
		name the simulator as WRENCH-Ext, along with the WRENCH simulator,
		which is created with original WRENCH.
		Finally, we describe experimental scenarios to evaluate our simulation model 
		using a standalone Python prototype, WRENCH and WRENCH-Ext simulator.

@@

            We separate our simulation model in two components, the IO
			Controller and the Memory Manager, which together simulate 
			file read/write mechanisms in Linux (Figure~\ref{fig:interaction}). 
			The Memory Manager simulates I/Os to cache, flushing, periodical flushing, 
			and cache eviction, whereas the IO Controller simulates file reads and writes. 			
			To read or write a chunk of file, a simulated application sends a request 
			to the IOController.
			Based on the memory status in Memory Manager, the IOController 
			orchestrates cache and disk reads/writes, flushing, cache eviction 
			to simulate I/O time.

			\begin{figure}
   				\centering
   				\includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
				   \caption{Overview of the simulated writeback page cache.
				   Application sends file read and file write requests to IOController, 
				   then IOController manages orchestrate flushing, eviction, cache and 
				   disk accesses to fulfill the requests with the MemoryManager object. 
				   Memory and Disk are responsible for simulating I/O time.}
				   \label{fig:interaction}
			\end{figure}

@@

			We leverage existing storage simulation models to simulate the memory,
			which is characterized with storage capacity, read and write bandwidths 
			and device latency. 
			These features are captured in simulation frameworks and applicable 
			to memory simulation \cite{lebre2015}. 
			In addition, leveraging existing disk models allows us 
			to take advantage of bandwidth sharing between concurrent disk I/O to 
			simulate parallel memory accesses.

			\paragraph*{Memory Manager}
			Memory Manager simulates two parallel threads presented in \ref{fig:interaction}.
			The first thread is the main thread, which is responsible for the simulation 
			of data flushing, cache eviction and I/Os to cache. 
			The second thread is a simulated background thread, in which Memory Manager 
			periodically searches and flushes expired dirty data in page cache LRU lists 
			to disks.

			\paragraph*{Page cache LRU lists}
			In Linux kernel, the page cache LRU lists contain file pages. However, 
			due to the large number of file pages, managing lists of pages may 
			require significant computation overhead. 
			Thus, we introduce the concept of a data block as a unit to represent data 
			cached in memory. A data block is a subset of file pages stored in
            page cache that were accessed in the same I/O operation. 
			A data block has information about file name, block size, last access 
			time, a dirty flag that represents whether the data is clean (0) 
			or dirty (1), and the time it is written to cache as dirty data. 
			Blocks can have different sizes and a given file can have multiple 
			data blocks in page cache. In addition, a data block can be split into an 
			arbitrary number of smaller blocks.

@@

			With the data block definition, page cache LRU lists can be modeled with 
			two lists of data blocks: an active list and an inactive list, ordered by 
			last access time (earliest first, see Figure~\ref{fig:lrulist}).
			As in the kernel, our simulator keeps the size of the active list under
			twice the size of the inactive list by moving least recently 
            used data blocks from the active list to the inactive list.
			Modeling page cache LRU lists as lists of data blocks reduces the
			overhead of the simulator, as data blocks are obviously less
			numerous than file pages, and at the same time achieves the accuracy in 
			I/O time simulation that can be seen in the later section.

			At a specific time, an arbitrary file can be partially cached, completely cached, 
			or uncached.
			Cached data is composed of data blocks existing in one or both
            of the page cache LRU lists. 
            Initially, blocks accessed for the first time are added to the inactive list in 
            page cache. 
            If the blocks residing the inactive list are subsequently accessed, 
            they will be moved to the top of the active list
            Cached blocks written to cache are marked as dirty blocks until 
            they are flushed to disk.

			\paragraph*{File read and write}
            Our model supports chunk-by-chunk file read and write simulation
            with a user-defined chunk size, in which chunks are read/written
            until file is entirely read/written.
            The simulated LRU lists are used as follows. 
            When a file is read, its data can be partially cached, completely cached, 
            or uncached. 
            In the scope of this study, our model support file accesses in chunk 
            but not file random access.
            Thus, we simplify the data read order, in which uncached data is read 
            from disk first, and cached data is only read after all uncached data is read 
            (left to right in Figure~\ref{fig:read_order}).
            When a chunk of uncached data is read, a new clean block is created 
            and put at the top of the inactive list. 
            For a chunk of cached data, one or more existing data blocks in two 
            page cache LRU lists are accessed again, and moved to the active list. 
			If the blocks are clean, we only need to keep the information about the 
			amount of data as they are available for eviction. 
			Thus, these re-accessed clean blocks can be merge into a bigger block 
			to reduce the number of block objects and put it in the active list. 
            If the blocks are dirty, since the blocks need to keep the information 
            of entry time for periodical flushing, they are only moved to the active list. 
			Because the chunk size and block size can be different, there are circumstances 
			in which a block is not entirely read. 
			In this case, the block is split into two smaller blocks, one is re-accessed 
			and the other remains the same.

			\begin{figure}
   				\centering
   				\includegraphics[width=\columnwidth]{figures/read_order.pdf}
   				\caption{File data read order from left to right, in which uncached data 
   				is read first, then cached data from inactive list and active list is read later.}	
   				\label{fig:read_order}
			\end{figure}

			For writes, we assume that all data to be written is 
			uncached and it is written to cache as dirty data. 
			Thus, each time a chunk is written, we create a block of dirty data 
			and put it in the inactive list.

@@

			\paragraph*{Flushing}
			Next, we simulate data flushing, which runs in the main simulated 
			thread and writes back dirty data to disk. 
			The data flushing simulation function traverses the sorted 
			inactive list first, and then the active list, looks for least recently used 
			dirty blocks, writes them to disk through the storage device model, 
			and sets dirty flags of these blocks to 0 until the total flushed amount 
			reaches the amount to be flushed or there is no dirty data left in cache. 
			Similar to file read and write, the size of flushed data and block size 
			can be different, there are cases that the last block is not entirely flushed.  
			In that case, the last block is split into two blocks, one that is flushed 
			and one that remains dirty.
			The flushing time is simulated with storage model.

			\paragraph*{Cache eviction}	
			Similarly to data flushing, our cache eviction simulation runs on 
			the main thread, frees up the page cache by traversing and deleting 
			least recently used clean data blocks in the inactive list.
			The amount of evicted data is passed in and data blocks are deleted 
			from the inactive list until total evicted data reaches the amount 
			passed in, or there is no clean block left in the list.
			If the last evicted block is not entirely evicted, the block is split, 
			and only one block is deleted.
			The cache eviction simulation does not add up to the simulated time 
			since cache eviction time is negligible in real systems.

			\begin{algorithm}\caption{Periodical flushing simulation}\label{alg:pdflush}
				\small
				\begin{algorithmic}[1]
					\Input
        				\Desc{in}{page cache inactive list}
        				\Desc{ac}{page cache active list}
						\Desc{t}{predefined flushing time interval}
						\Desc{exp}{predefined expiration time}						
						\Desc{sm}{storage simulation model}
   					\EndInput
   					\While{host is on}
						\State blocks = expired\_blocks(exp, in) + expired\_blocks(exp, ac) 
						\State flushing\_time = sm.write(blocks)
						\State flag blocks as clean data
						\If{flushing\_time $<$ t}
							\State sleep(t - flushing\_time)
						\EndIf
					\EndWhile
				\end{algorithmic}
			\end{algorithm}

			\paragraph*{Periodical flushing}	
			The periodical flushing simulation is run on a simulated background thread 
			to flush expired dirty data blocks from cache to disk.
			Similar to expired dirty pages in real Linux systems, a dirty block in our model 
			is considered expired if: 
			\begin{enumerate}
				\item It is dirty and 
				\item The time from its entry time to current time is longer than the 
				predefined expiration time. 
			\end{enumerate}
			The simulation algorithm of periodical flushing is an infinite loop in which 
			Memory Manager searches and flushes these expired dirty blocks to disk, 
			as described in Algorithm~\ref{alg:pdflush}. 
			The loop ends when the host is off (line 7). 
			In each repetition, Memory Manager finds expired dirty blocks in two 
			page cache LRU lists (line 8), simulates writes of data of these blocks 
			to disk (line 9), and mark them as clean (line 10).
			If the flushing time does not exceed our time interval, the thread is put 
			to sleep for the remaining time (lines 11-13). 
			Then based on the state of the host at current simulated time, 
			the algorithm continues or finishes the loop.
			Because periodical flushing is simulated as a background thread, it can be 
			concurrent with disk I/O on main thread. This parallelism is handled by the 
			storage model and reflected in simulated I/O time.

@@

			As mentioned, in our model, file reads and writes can be completed by 
			reading or writing chunk-by-chunk until entire file is read or written.
